Foreign direct investment, corporate social responsibility, and malaria control in Mozambique - trends, risks, and opportunities

### 2017-07-11 paper.Rmd
big update

@@

# Clean up columns
db <- data.frame(db)
macro_names <- names(db)
for (j in 1:length(macro_names)){
  number <- j
  while(grepl('Var', macro_names[number])){
    number <- number -1
  }
  macro_names[j] <- macro_names[number]
}
micro_names <- as.character(db[1,])
combined_names <- gsub(' ', '', paste0(macro_names, '..', ifelse(is.na(micro_names), '', micro_names)))
# Remove extra columns and replace column names
db <- db[-1,]
names(db) <- combined_names

>>>>

### Sources of FDI in Mozambique

```{r}
srcs <- readxl::read_excel('data/data.imf.orgSLASHregular.aspx?key=60564262.xls', skip = 4)
# Rename columns
srcs <- srcs[,1:3]
srcs <- srcs[2:nrow(srcs),]
names(srcs) <- c('from_country', 'inward', 'outward')
# Make numberic
srcs <- data.frame(srcs)
for (j in 2:3){
  this_column <- as.numeric(as.character(gsub('.', '', srcs[,j], fixed = TRUE)))
  srcs[,j] <- this_column
}
srcs <- srcs %>%
  filter(from_country != 'World')

# Get amount by inward
x <- srcs %>%
  arrange(desc(inward)) %>%
  filter(!is.na(inward) & inward > 0) %>%
  mutate(from_country = ifelse(from_country == 'China, P.R.: Mainland',
                               'China',
                               ifelse(from_country == 'United Arab Emirates',
                                      'U.A.E.',
                                      ifelse(from_country == 'United Kingdom',
                                             'U.K.',
                                             from_country)))) %>%
  mutate(from_country = factor(from_country, levels = from_country))
```

According to the IMF, `r nrow(x)` countries had foreign direct investment in Mozambique as of the end of 2015. The below chart shows the top 10 countries ranked by amount invested in Mozambique.

```{r}
ggplot(data = x[1:10,],
       aes(x = from_country,
           y = inward)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           fill = 'darkblue') +
  labs(x = 'Source of investment',
       y = '2015 USD, Millions',
       title = 'Sources of FDI in Mozambique',
       subtitle = 'By investor country') +
  theme_bw() +
    theme(axis.text.x = element_text(angle = 90)) 
```

What stands out is the presence of many "non-Western" countries in the top 10. Though data are not available over time (ie, we cannot make temporal comparisons), the landscape of investment 20 years prior would have been both (a) far smaller and (b) primarily composed of western countries.

@@

# Clean up columns
db <- data.frame(db)
macro_names <- names(db)
macro_names <- ifelse(grepl('X__', macro_names), 'Var', macro_names)

for (j in 1:length(macro_names)){
  number <- j
  while(grepl('Var', macro_names[number])){
    number <- number -1
  }
  macro_names[j] <- macro_names[number]
}
micro_names <- as.character(db[1,])
combined_names <- gsub(' ', '', paste0(macro_names, '..', ifelse(is.na(micro_names), '', micro_names)))
# Remove extra columns and replace column names
db <- db[-1,]
names(db) <- combined_names

### 2017-07-11 paper.tex
big update

@@

@@

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1} \end{center}

@@

@@

@@

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-15-1} \end{center}

>>>>

\subsection{Abstract}\label{abstract}

Foreign direct investment (FDI) in Mozambique has increased rapidly in
the last two decades. The growing interest in corporate social
responsibility (CSR) - combined with a recent push for malaria
elimination - suggest the need for a critical examination of the role of
FDI in malaria control and elimination. Through a systematic review of
the literature, we find that there has been a notable increase in
research pertaining to FDI, CSR and malaria in recent years. Consensus
among researchers is that this increase has not been accompanied by a
substantial private sector investment in malaria control (with a few
notable exceptions). This suggests a potential opportunity for scaling
up privately driven malaria control as a means to both improve public
health and increase return on investment. However, given the lack of
coordination between public and private sectors, even when interests
coincide, an over-reliance on foreign and private initiative for funding
eradication is not without risks.

@@

\subsubsection{Sources of FDI in
Mozambique}\label{sources-of-fdi-in-mozambique}

According to the IMF, 69 countries had foreign direct investment in
Mozambique as of the end of 2015. The below chart shows the top 10
countries ranked by amount invested in Mozambique.

What stands out is the presence of many ``non-Western'' countries in the
top 10. Though data are not available over time (ie, we cannot make
temporal comparisons), the landscape of investment 20 years prior would
have been both (a) far smaller and (b) primarily composed of western
countries.

@@

@@

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1} \end{center}

@@

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-14-1} \end{center}

@@

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-17-1} \end{center}
